```{r setup, results=FALSE, message = FALSE}
options(width = 120)
library(data.table)
library(hash)
library(plotly)
library(arules)
```

### Reading data:

```{r message = FALSE}
cards = fread(cmd = '7z e -so "jan_2019_games_arenas_1_to_13.7z"', header=TRUE)
```

### Transforming decks to lists and adding winner_deck information:

```{r}
cards$player_deck_list <- strsplit(cards$player_deck, "_")
cards$opp_deck_list <- strsplit(cards$opp_deck, "_")
cards[, winner_deck_list := ifelse(has_won, player_deck_list, opp_deck_list)]
head(cards, 1000)
```

### Util function for apriori algorithm- takes frequent itemsets of size N and generates candidate itemsets of size N+1:

```{r message = FALSE}
createCandidates <- function(prevCandidates) {
  if (length(prevCandidates) == 0) {
    return(list())
  }
  n <- length(prevCandidates[[1]])
  
  result <- list()
  
  for (i in 1:length(prevCandidates)) {
    for (j in i:length(prevCandidates)) {
      if ((n == 1 || identical(prevCandidates[[i]][1:(n-1)], prevCandidates[[j]][1:(n-1)])) &&
          as.character(prevCandidates[[i]][n]) != as.character(prevCandidates[[j]][n])) {
        new_list <- c(prevCandidates[[i]], prevCandidates[[j]][n])
        result[[length(result)+1]] <- new_list
      }
      else if (n != 1 && !identical(prevCandidates[[i]][1:(n-1)], prevCandidates[[j]][1:(n-1)])) {
        break
      }
    }
  }
  
  result
}

```

### Apriori algorithm (hybrid solution- one approach is faster and more memory-efficient for smaller sizes, one for bigger sizes (and smaller amount of candidates)):

```{r message = FALSE}
DECK_SIZE=8
HYBRID_SOLUTION_SIZE_THRESHOLD = 4 # the solution is too demanding in terms of time without hybrid solution
apriori <- function(df, support) {
  n <- 2 * nrow(df)
  player_deck_counts <- table(unlist(c(df$player_deck_list, df$opp_deck_list)))
  winner_deck_counts <- table(unlist(df$winner_deck_list))
  player_deck_counts_filtered <- player_deck_counts[player_deck_counts >= n * support]
  curFreqSet <- lapply(names(player_deck_counts_filtered), function(card) list(card))
  res <- mapply(function(x, y) c(itemset = list(x), pick_ratio = y / n, win_ratio = ifelse(is.na(winner_deck_counts[x][[1]]), 0, winner_deck_counts[x][[1]]) / y, size=1, name=x), names(player_deck_counts_filtered), as.list(player_deck_counts_filtered), SIMPLIFY = FALSE)
  for (size in 2:HYBRID_SOLUTION_SIZE_THRESHOLD) {
    cands = createCandidates(curFreqSet)
    cand_counts <- hash()
    winner_counts <- hash()
    for (cand in cands) {
      name <- paste(cand, collapse=", ")
      cand_counts[[name]] <- 0
      winner_counts[[name]] <- 0
    }
    lapply(combn(1:DECK_SIZE, size, simplify = FALSE), function(comb) {
      player_counts_on_pos <- table(sapply(df$player_deck_list, function(lst) paste(lst[comb], collapse = ", ")))
      opp_counts_on_pos <- table(sapply(df$opp_deck_list, function(lst) paste(lst[comb], collapse = ", ")))
      winner_counts_on_pos <- table(sapply(df$winner_deck_list, function(lst) paste(lst[comb], collapse = ", ")))
      lapply(names(player_counts_on_pos), function(name) {
        if(exists(name, cand_counts)) {
          cand_counts[[name]] <- cand_counts[[name]] + player_counts_on_pos[[name]]
        }
      })
      lapply(names(opp_counts_on_pos), function(name) {
        if(exists(name, cand_counts)) {
          cand_counts[[name]] <- cand_counts[[name]] + opp_counts_on_pos[[name]]
        }
      })
      lapply(names(winner_counts_on_pos), function(name) {
        if(exists(name, winner_counts)) {
          winner_counts[[name]] <- winner_counts[[name]] + winner_counts_on_pos[[name]]
        }
      })
    })
    curFreqSet <- lapply(names(cand_counts), function(name) {
      picks <- cand_counts[[name]]
      wins <- if (exists(name, winner_counts)) winner_counts[[name]] else 0
      if (picks >= n * support) {
        lst <- unlist(strsplit(name, ", "))
        newSet <- list(itemset = lst, pick_ratio = picks / n, win_ratio = wins / picks, size = length(lst), name=name)
        return(newSet)
      }
      else {
        return(NULL)
      }
    })
    curFreqSet <- curFreqSet[lengths(curFreqSet) > 0]
    res <- c(res, curFreqSet)
    curFreqSet <- lapply(curFreqSet, function(x) x$itemset)
  }
  for (size in (HYBRID_SOLUTION_SIZE_THRESHOLD+1):DECK_SIZE) {
    cands = createCandidates(curFreqSet)
    curFreqSet <- lapply(cands, function(lst) {
      rows <- sum(colSums(sapply(df$player_deck_list, function(deck) lst %in% deck)) == length(lst)) + sum(colSums(sapply(df$opp_deck_list, function(deck) lst %in% deck)) == length(lst))
      if (rows >= support * n) {
        rows_won <- sum(colSums(sapply(df$winner_deck_list, function(deck) lst %in% deck)) == length(lst))
        return(list(itemset = lst, pick_ratio = rows / n, win_ratio = rows_won / rows, size = length(lst), name=paste(lst, collapse=", ")))
      } else {
        return(NULL)
      }
    })
    curFreqSet <- curFreqSet[lengths(curFreqSet) > 0]
    res <- c(res, curFreqSet)
    curFreqSet <- lapply(curFreqSet, function(x) x$itemset)
  }
  res
}
```

```{r message = FALSE}
aprioriToDt <- function(aprioriRes) {
  res <- rbindlist(lapply(aprioriRes, function(el) {
    el$itemset <- el$name
    el
  }))
  res[, itemset:= sapply(res$name, function(name) strsplit(name, ", "))]   # rbindlist doesn't transform non-atomic values
  res
}
```

### Faster implementation- using external library 'arules':

```{r message = FALSE}
findFreqItems <- function(df, support) {
  transactions <- as(c(df$player_deck_list, df$opp_deck_list), "transactions")
  frequent_itemsets <- eclat(transactions, parameter = list(support = support))
  res <- data.table(as(frequent_itemsets, "data.frame"))
  res$itemset <- lapply(strsplit(gsub("[{}]", "", res$items), ",\\s*"), as.character)
  res$pick_ratio <- res$support
  res$size <- sapply(res$itemset, function(items) length(items))
  res$name <- sapply(res$itemset, function(items) paste(items, collapse = ", "))
  res$items <- NULL
  res$support <- NULL
  res$count <- NULL
  
  wins_transactions <- as(df$winner_deck_list, "transactions")
  winning_itemsets <- eclat(wins_transactions, parameter = list(support = 0.001))
  winners <- data.table(as(winning_itemsets, "data.frame"))
  winners$name <- sapply(lapply(strsplit(gsub("[{}]", "", winners$items), ",\\s*"), as.character), function(items) paste(items, collapse = ", "))
  winners$wins <- winners$count
  winners$items <- NULL
  winners$support <- NULL
  winners$count <- NULL

  res_with_wins <- merge(res, winners, by = "name", all.x = TRUE)
  res_with_wins$wins <- ifelse(is.na(res_with_wins$wins), 0, res_with_wins$wins)
  n <- 2 * nrow(df)
  res_with_wins$win_ratio <- res_with_wins$wins / res_with_wins$pick_ratio / n
  res_with_wins
}
```



```{r message = FALSE, results = FALSE}
SUPPORT = 0.01
freqItemsDt = findFreqItems(cards, SUPPORT)
```
```{r}
head(freqItemsDt, 1000)
```

### Util functions to generate plots and one function to take 100 most frequent itemsets for each size:

```{r message = FALSE}
createTop100 <- function(freqItemsDt, df) {
  n <- 2 * nrow(df)
  top100 <- freqItemsDt[order(-pick_ratio), .SD[1:100], by=size]
  top100 <- top100[pick_ratio > 0] # filtering NA's
  top100
}

genPickRatioPlot <- function(df) {
  plot_ly(df, x=~factor(name,levels = unique(name[order(size, -pick_ratio)])), y=~pick_ratio, color=~factor(size, levels = unique(size)), hovertemplate="Itemset: %{x}<br>Pick Ratio: %{y}<br>", type="bar") %>% layout(xaxis = list(title="Itemset", showticklabels = FALSE), yaxis = list(title="Pick Ratio")) %>% layout(title="Pick Ratio by Itemset")
}

genAnimatedPickRatioPlot <- function(df, frame) {
  plot_ly(df, x=~factor(name,levels = unique(name[order(size, -pick_ratio)])), y=~pick_ratio, color=~factor(size, levels = unique(size)), frame=frame, hovertemplate="Itemset: %{x}<br>Pick Ratio: %{y}<br>", type="bar") %>% layout(xaxis = list(title="Itemset", showticklabels = FALSE), yaxis = list(title="Pick Ratio")) %>% layout(title="Pick Ratio by Itemset")  %>%
    animation_opts(
      500, redraw = FALSE
    )
}

genWinRatioPlot <- function(df) {
  plot_ly(df, x=~factor(name,levels = unique(name[order(size, -win_ratio)])), y=~win_ratio, color=~factor(size, levels = unique(size)), hovertemplate="Itemset: %{x}<br>Win Ratio: %{y}<br>", type="bar") %>% layout(xaxis = list(title="Itemset", showticklabels = FALSE), yaxis = list(title="Win Ratio")) %>% layout(title="Win Ratio by Itemset")
}

genAnimatedWinRatioPlot <- function(df, frame) {
  plot_ly(df, x=~factor(name,levels = unique(name[order(size, -win_ratio)])), y=~win_ratio, color=~factor(size, levels = unique(size)), frame=frame, hovertemplate="Itemset: %{x}<br>Win Ratio: %{y}<br>", type="bar") %>% layout(xaxis = list(title="Itemset", showticklabels = FALSE), yaxis = list(title="Win Ratio")) %>% layout(title="Win Ratio by Itemset")  %>%
    animation_opts(
      500, redraw = FALSE
    )
}
```

```{r message = FALSE}
top100 <- createTop100(freqItemsDt, cards)
```

### Pick ratio/Win ratio based on arena:

```{r}
genPickRatioPlot(top100)
```
```{r}
genWinRatioPlot(top100)
```
```{r message = FALSE, results = FALSE}
freqItemsByArena <- lapply(unique(cards$arena_id), function(arenaId) {
  fromArena <- cards[arena_id == arenaId]
  freqFromArenaDt <- findFreqItems(fromArena, SUPPORT)
  freqFromArenaDt[, arena_id := arenaId]
  top100Arenas <- createTop100(freqFromArenaDt, fromArena)
  top100Arenas
})
freqItemsByArena <- rbindlist(freqItemsByArena)
```
```{r}
genAnimatedPickRatioPlot(freqItemsByArena, ~arena_id)
```
```{r}
genAnimatedWinRatioPlot(freqItemsByArena, ~arena_id)
```

### Pick ratio/Win ratio based on day:

```{r message = FALSE, results = FALSE}
cards$day <- as.Date(cards$timestamp)
freqItemsByDay <- lapply(unique(cards$day), function(dayId) {
  fromDay <- cards[day == dayId]
  freqFromDayDt <- findFreqItems(fromDay, SUPPORT)
  freqFromDayDt[, day := as.character(dayId)]
  top100Days <- createTop100(freqFromDayDt, fromDay)
  top100Days
})
freqItemsByDay <- rbindlist(freqItemsByDay)
```
```{r}
genAnimatedPickRatioPlot(freqItemsByDay, ~day)
```
```{r}
genAnimatedWinRatioPlot(freqItemsByDay, ~day)
```

### Finding associactions between cards:

```{r}
findAssociacions <- function(cur_itemset, candidates, support) {
  item_pick_ratio = cur_itemset$pick_ratio
  max_pick_ratio = item_pick_ratio / support # support(cand => itemset) = item_pick_ratio / cand_pick_ratio >= support
  candidates = candidates[pick_ratio <= max_pick_ratio]
  cur_items <- cur_itemset$itemset
  candidates_subsets <- candidates[sapply(itemset, function(items) all(items %in% cur_items) && !all(cur_items %in% items))]
  mapply(function(pick_ratio, name, items) list(left_side_pick_ratio=pick_ratio, left_side=name, right_side=toString(setdiff(cur_items, items)), right_side_support=item_pick_ratio / pick_ratio), candidates_subsets$pick_ratio, candidates_subsets$name, candidates_subsets$itemset, SIMPLIFY = FALSE)
}
x <- apply(top100, MARGIN=1, function(itemset) findAssociacions(itemset, freqItemsDt, 0.33))
result <- list()
for (res_list in x) {
  result <- c(result, res_list)
}
result <- rbindlist(result)
head(result[order(right_side_support, decreasing = TRUE)], 1000)
```

```{r}
hover_text <- paste(
  "<b>Left Side:</b> ", result$left_side, "<br>",
  "<b>Left Side Pick Ratio:</b> ", result$left_side_pick_ratio, "<br>",
  "<b>Right Side:</b> ", result$right_side, "<br>",
  "<b>Right Side Support:</b> ", result$right_side_support, "<br>"
)

plot_ly(result, x = ~left_side_pick_ratio, y = ~right_side_support, type = "scatter", mode = "markers", 
        hovertemplate = hover_text, marker = list(size = 10, color = "blue"))

```

### Creating recommendation table- it stores information about how many times each user played each card in his history:

```{r}
exploded <- cards[, .(tag, card = unlist(player_deck_list)), by = seq_len(nrow(cards))]
usage <- dcast(exploded, tag ~ card, fun.aggregate = length)
head(usage)
```

Performing SVD decomposition on recommendation table:
  
```{r message = FALSE}
usage_matrix <- as.matrix(usage[, -1])

svd_res <- svd(usage_matrix)

U <- svd_res$u
S <- diag(svd_res$d)
V <- svd_res$v
```

Grouping players based on the most dominant concept:
  
```{r message = FALSE}
K = ceiling(nrow(S) / 2)
U <- U[1:K, ]
S <- S[1:K, 1:K]
V <- V[, 1:K]
groups <- max.col(apply(U, c(1,2), abs))
user_concept <- head(data.table(userId = usage$tag, cluster_number = groups), 1000)
```
```{r}
plot_ly(user_concept, x=~userId, y=~cluster_number, type = "scatter", color=~cluster_number) %>%
  layout(xaxis = list(categoryorder = "total ascending"))
```

### Recommendation system: based on provided 4 cards- we calculate preferences regarding the concepts, which then allow us to calculate cards preferences; then we need to remove our 4 cards from the list and take 4 with the highest values

```{r}
cards_number = nrow(V)
create_remcommendations <- function(chosen_cards) {
  chosen_cards_idxs <- sapply(chosen_cards, function(card) which(names(usage) == card))
  preferences <- rep(0, cards_number)
  preferences[chosen_cards_idxs] <- 1
  concept_preferences <- preferences %*% V
  preference_prediction <- concept_preferences %*% t(V)
  preference_prediction <- apply(preference_prediction, c(1,2), abs)
  names(preference_prediction) <- colnames(usage)[-1] # cut 'tag' colname
  setdiff(names(sort(preference_prediction, decreasing = TRUE)), chosen_cards)[1:4]
}
create_remcommendations(c("balloon", "baby-dragon", "archers", "arrows"))
```
